### 2019-01-22 diagram_NAR.tex
First commit

%\documentclass[tikz]{standalone}

\begin{tikzpicture}[node distance=2cm]

\node (react2) [startend] {Condition 2};
\node (react1) [startend, left = 10pt of react2] {Condition1};
\node (rsup) [startend, right = 10pt of react2] {...};

\node (suu)[above=5pt of react2,IO]{\relsize{+1}\sf Reactivities};

\node(sampling)[Bprocess, below = 20pt of react2]{Stochastic Structural Sampling};
\node (seq)[startend, left =1cm of sampling]{RNA sequence};

\node[IO]  at (seq|-react1) {\sf\bfseries \relsize{+2}Input};

\draw  [arrow] (seq) to (sampling);

\node (sample2) [startend, below= 15pt of sampling] {Sample 2};
\node (sample1) [startend, left = 10pt of sample2] {Sample 1};
\node (samplex) [startend, right = 10pt of sample2] {...};

\node(Merge)[Vprocess, below = 15pt of sample2, ]{{Merge samples}};

\node(clustering)[Bprocess, below = 15pt of Merge]{Clustering with k clusters};
\draw[arrow](Merge) -- (clustering);
\node(dec1) [decision, below=15pt of clustering]{Stopping \\ Criterion};
\draw[arrow] (clustering) to (dec1);
\node(optimaln)[startend, below of = dec1]{Centroid Structures};

\node(n+)[VSprocess, left of = dec1,xshift=-3cm]{\rotatebox{90}{k++}};
\draw[arrow] (n+)  |- (clustering);
\node(dec2) [decision, below=15pt of optimaln]{Pareto\\Optimal?};
\draw[arrow](optimaln) to (dec2);
\node(nnoptim)[VSprocess,left of = dec2, xshift=-3cm ]{\rotatebox{90}{Reject centroid}};

\node(optimalcentroid)[startend, below of = dec2]{Optimal centroid(s)};

\node(output)[right of =optimalcentroid, IO] (outcap) {\sf\bfseries \relsize{+2}Output};
\node(vert)[right of = optimalcentroid, xshift=1cm ]{};

\draw[arrow] (dec1) -- node[anchor=north,arrowcaption] {no} (n+);
\draw[arrow] (dec1) -- node [anchor=west,arrowcaption]{yes} (optimaln);

\draw[arrow] (dec2) -- node[anchor=north,arrowcaption] {no} (nnoptim);
\draw[arrow] (dec2) -- node [anchor=west,arrowcaption]{yes} (optimalcentroid);

%\draw[arrow](Cbz) |- (dec2);

\draw  [arrow] (react1) -- (sampling.north-|react1);
\draw  [arrow] (react2)-- (sampling.north-|react2);
\draw  [arrow] (rsup) -- (sampling.north-|rsup);

\draw[arrow] (sampling.south-|sample1) -- (sample1);
\draw[arrow] (sampling.south-|sample2) -- (sample2);
\draw[arrow] (sampling.south-|samplex) -- (samplex);

\draw[arrow] (sample1) -- (Merge.north-|sample1);
\draw[arrow] (sample2) -- (Merge.north-|sample2);
\draw[arrow] (samplex) -- (Merge.north-|samplex);
%\draw[arrow] (sample1) -- ++(-3cm,0);
%\draw[arrow] (sample2) -- ++(-3cm,0);
%\draw[arrow] (samplei) -- ++(-3cm,0);
%\draw[arrow] (samplex) -- ++(-3cm,0);

\begin{pgfonlayer}{background}
\node[fit=(seq.west|-react1) (react1) (react2) (rsup) (suu), input] (bginput1) {};
\node[fit=(seq)(seq.west|-react1), input] (bginput2) {};
\node[fit=(outcap) (optimalcentroid), output] (bgoutput) {};
\end{pgfonlayer}

\end{tikzpicture}
%\end{document}