### 2019-06-19 SuppMat.tex

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage{float}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
\usepackage{adjustbox}   
%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%%%%%%%%%%%%%%%%%%%%%%%%%formatting cells colors
\usepackage[table,x11names,dvipsnames]{xcolor}
 \usepackage{collcell}
 \usepackage{array}
 \usepackage{tikz}
 \usepackage{relsize}
  \usepackage{pgfkeys}
 \usepackage{graphicx}
 \usepackage{xspace}
 \usepackage{amsmath,amssymb}

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information
\usepackage{multicol}
\usepackage{booktabs}
\usepackage{todonotes}
\usepackage{inconsolata}

\input{macros.tex}

\title{\OurTool{}:  Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles\\[1em]
	Supplementary material
}
\author{}
\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 
\newcommand{\Start}{s}
\newcommand{\End}{e}

\begin{document}
\maketitle

\section{Detailed results on Cordero~\emph{et al} dataset}
\subsection{Structure prediction ordered by RNA}

 \Heatset{min=0.4,  
	max=1,   
	max colour=Aquamarine3, % colour at maximum
	min colour=BadCol,      % colour at minimum
	Min colour=BadCol, % colour for values below min
	Max colour=Aquamarine3   % colour for values above max
}
\begin{table}[H]
\centering
\newcommand{\B}{\multicolumn{1}{c}{$\bullet$}}
\newcommand{\BR}{\multicolumn{1}{c}{$\bullet$}}
\newcommand{\D}{\multicolumn{1}{c}{\color{gray}$\circ$}}
\newcommand{\DR}{\multicolumn{1}{c}{\color{gray}$\circ$}}
\begin{adjustbox}{max width=1\textwidth}
 \begin{tabular}{@{}lHHHHHHHHHHHHHHHH@{}}
 	\toprule
 	                              & \multicolumn{8}{l}{ \OurTool{}}                                     & \multicolumn{4}{l}{\Software{RNAfold} -- MFE mode} & \multicolumn{4}{l}{\Software{RNAfold} -- MEA mode}  \\
 	\textbf{CMCT}                 & \D    & \B    & \D       & \D    & \D    & \B       & \B    & \BR   & \D       & \D    & \B    & \DR                     & \D    & \D    & \B       & \D                       \\
 	\textbf{NMIA}                 & \D    & \D    & \B       & \D    & \B    & \D       & \B    & \BR   & \D       & \D    & \D    & \BR                     & \D    & \D    & \D       & \B                       \\
 	\textbf{DMS}                  & \D    & \D    & \D       & \B    & \B    & \B       & \D    & \BR   & \D       & \B    & \D    & \DR                     & \D    & \B    & \D       & \D                       \\ \midrule
 	5SRNA,E.Coli                  & 0.25  & 0.238 & 0.247    & 0.244 & 0.247 & 0.247    & 0.254 & 0.241 & 0.241    & 0.686 & 0.254 & 0.686                   & 0.241 & 0.686 & 0.269    & 0.686                    \\
 	glycineriboswitch,F.nucleatum & 0.568 & 0.627 & 0.868    & 0.952 & 0.868 & 0.658    & 0.868 & 0.868 & 0.306    & 0.313 & 0.395 & 0.313                   & 0.593 & 0.593 & 0.693    & 0.6                      \\
 	cidGMPriboswitch,V.Cholerae   & 0.654 & 0.77  & 0.654    & 0.77  & 0.654 & 0.77     & 0.654 & 0.77  & 0.77     & 0.77  & 0.667 & 0.77                    & 0.77  & 0.72  & 0.667    & 0.77                     \\
 	P4P6domain                    & 0.864 & 0.856 & 0.881    & 0.864 & 0.864 & 0.856    & 0.864 & 0.856 & 0.837    & 0.808 & 0.773 & 0.714                   & 0.845 & 0.808 & 0.781    & 0.722                    \\
 	adenineriboswitch,add         & 1     & 0.956 & 1        & 0.977 & 1     & 0.956    & 1     & 1     & 1        & 0.333 & 0.41  & 1                       & 1     & 0.333 & 0.356    & 0.306                    \\
 	tRNAphenylalanineyeast        & 0.286 & 0.746 & 0.976    & 0.746 & 0.976 & 0.746    & 0.976 & 0.976 & 0.976    & 0.976 & 0.976 & 0.976                   & 0.334 & 0.976 & 0.976    & 0.976                    \\ \midrule
 	Average                       & 0.604 & 0.7   & 0.77     & 0.76  & 0.77  & 0.705    & 0.77  & 0.785 & 0.69     & 0.65  & 0.58  & 0.74                    & 0.63  & 0.69  & 0.62     & 0.68                     \\ \bottomrule
 \end{tabular}
\end{adjustbox}
\caption{GM of the structures predicted with \OurTool{}, compared to predictions of \Software{RNAfold} in energy mimization (MFE) and accuracy maximization (MEA) modes.}
%( ref. $supplementary\_data/benchmark\_6RNAs$)}
\end{table}
\subsection{Reproducibility}
\begin{table}[H]
\centering
\begin{adjustbox}{max width=1\textwidth}
 \begin{tabular}{lHHHHHHHHHHHHHHHHHHHHH}\toprule
 &\multicolumn{3}{l}{ NMIA} & \multicolumn{3}{l}{ DMS} & \multicolumn{3}{l}{CMCT} &\multicolumn{3}{l}{ NMIA+DMS} & \multicolumn{3}{l}{ NMIA+CMCT} & \multicolumn{3}{l}{DMS+CMCT}& \multicolumn{3}{l}{NMIA+DMS+CMCT} \\
% & \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM }& \multicolumn{3}{l}{ GM }  & \multicolumn{3}{l}{ GM } & \multicolumn{3}{l}{ GM } &\multicolumn{3}{l}{ GM }\\

Replicates& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3} & \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3} & \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3} & \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}& \multicolumn{1}{l}{Run 1} & \multicolumn{1}{l}{Run 2} & \multicolumn{1}{l}{Run 3}\\
\midrule
5SRNA,E.Coli	&	0.247	&	0.25	&	0.247	&	0.254	&	0.244	&	0.244	&	0.238	&	0.238	&	0.238	&	0.247	&	0.244	&	0.247	&	0.257	&	0.238	&	0.238	&	0.247	&	0.247	&	0.244	&	0.241	&	0.241	&	0.241	\\
glycineriboswitch,F.nucleatum	&	0.868	&	0.868	&	0.868	&	0.952	&	0.952	&	0.952	&	0.868	&	0.634	&	0.627	&	0.868	&	0.868	&	0.868	&	0.868	&	0.868	&	0.868	&	0.667	&	0.658	&	0.658	&	0.868	&	0.868	&	0.868	\\
cidGMPriboswitch,V.Cholerae	&	0.631	&	0.631	&	0.631	&	0.77	&	0.77	&	0.77	&	0.667	&	0.77	&	0.77	&	0.631	&	0.631	&	0.756	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	&	0.77	\\
P4P6domain	&	0.864	&	0.881	&	0.881	&	0.864	&	0.864	&	0.856	&	0.856	&	0.856	&	0.856	&	0.864	&	0.864	&	0.864	&	0.864	&	0.864	&	0.864	&	0.856	&	0.856	&	0.856	&	0.856	&	0.856	&	0.856	\\
adenineriboswitch,ad	&	1	&	1	&	1	&	0.956	&	0.956	&	0.956	&	1	&	0.956	&	0.956	&	1	&	1	&	1	&	1	&	1	&	1	&	0.956	&	0.956	&	0.956	&	1	&	1	&	1	\\
tRNAphenylalanineyeast	&	0.953	&	0.953	&	0.976	&	0.976	&	0.746	&	0.976	&	0.746	&	0.746	&	0.746	&	0.976	&	0.976	&	0.976	&	0.976	&	0.976	&	0.97	&	0.746	&	0.746	&	0.746	&	0.746	&	0.976	&	0.976	\\

\bottomrule
\end{tabular}
\end{adjustbox}
\caption{ GM of the predicted structures with \OurTool{} over 3 runs, in the presence of 1,2 and 3 sources of probing data}
%( ref. $supplementary\_data/benchmark\_6RNAs$)
\end{table}

\newpage

%\section{IPANEMAP Vs Rsample}
\section{Detailed results on Hadjin~\emph{et al} dataset}

 \begin{table}[h]

{\centering
\begin{adjustbox}{max width=1\textwidth}
 \begin{tabular}{@{}llll@{}}
 	\toprule
 	RNA ID & Description                      \\ \midrule
 	A      & PreQ1riboswitchBsubtilis         & 	M      & SARScoronaviruspseudoknot        \\
 	B      & 5domainof16SrRNAEcoli            & 	N      & 5SrRNAEcoli                      \\
 	C      & SignalrecognitonparticleRNAhuman & 	O      & cyclicdiGMPriboswitchVcholerae   \\
 	D      & GroupIintronAzoarcussp           & 	P      & 5domainof23SrRNAEcoli            \\
 	E      & HIV15primepseudoknotdomain       & 	Q      & RNasePBsubtilis                  \\
 	F      & Telomerasepseudoknothuman        & 	R      & GroupIIntronTthermophila         \\
 	G      & SAMIriboswitchTtengcongensis     & 	S      & HepatitisCvirusIRESdomain        \\
 	H      & tRNA phe Ecoli                   & 	T      & 5domainof16SrRNAHvolcanii        \\
 	I      & P546domainbI3groupIintron        & 	U      & GroupIIintronOiheyensis          \\
 	J      & TPPriboswitchEcoli               & 	V      & tRNAaspyeast                     \\
 	K      & AdenineriboswitchVvulnificus     & 	W      & LysineriboswitchTmaritime        \\
 	L      & FluorideriboswitchPsyringae      & 	X      & MBoxriboswitchBsubtilis          \\
 	\bottomrule
 \end{tabular}
\end{adjustbox}\\}

\caption{Labels used for the 24 RNAs in the Hadjin \emph{ et al.} dataset~\cite{Hajdin2013}.}
\label{clyu}

\end{table}

\begin{table}[H]
	\begin{adjustbox}{max width=1\textwidth,max totalheight=\textheight,keepaspectratio}
		\begin{tabular}{llHHHHHHHHHHHlHH}
			\toprule
			\textbf{RNA} & Length &  \multicolumn1c{Run1} &\multicolumn1c{Run2}&\multicolumn1c{Run3}&\multicolumn1c{Run4}&\multicolumn1c{Run5}&\multicolumn1c{Run6}&\multicolumn1c{Run7}&\multicolumn1c{Run8}&\multicolumn1c{Run9}&\multicolumn1c{Run10}& \multicolumn1c{Average}& \multicolumn1c{Sd}&\multicolumn1c{\textbf{MFE-RNAfold}} & \multicolumn1c{\textbf{MEA-RNAfold}}\\
			\midrule

			A	&	34	&	0.791	&	0.791	&	0.791	&		0.791	&	0.791	&	0.791	&	0.791	&	0.791	&	0.791	&	0.722	&	0.7841	&	0.0207	&	0.791	&	0.791	\\
			B	&	530	&	0.851	&	0.851	&	0.851	&		0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0.851	&	0	&	0.851	&	0.849	\\
			C	&	301	&	0.6	&	0.6	&	0.6	&		0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0.6	&	0	&	0.6	&	0.597	\\
			D	&	214	&	0.789	&	0.789	&	0.789	&		0.781	&	0.781	&	0.781	&	0.781	&	0.781	&	0.781	&	0.781	&	0.7834	&	0.0037	&	0.768	&	0.761	\\
			E	&	500	&	0.539	&	0.539	&	0.539	&		0.539	&	0.539	&	0.539	&	0.539	&	0.535	&	0.541	&	0.515	&	0.5364	&	0.0073	&	0.562	&	0.524	\\
			F	&	47	&	0.671	&	0.671	&	0.671	&		0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0.671	&	0	&	0.671	&	0.671	\\
			G	&	118	&	0.801	&	0.801	&	0.801	&		0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0.801	&	0	&	0.892	&	0.879	\\
			H	&	76	&	1	&	1	&	1	&		1	&	1	&	1	&	1	&	0.762	&	0.762	&	0.781	&	0.9305	&	0.1063	&	1	&	1	\\
			I	&	155	&	0.982	&	0.982	&	0.982	&		0.982	&	0.982	&	0.982	&	0.982	&	0.982	&	0.982	&	0.964	&	0.9802	&	0.0054	&	0.964	&	0.982	\\
			J	&	79	&	0.914	&	0.914	&	0.914	&		0.914	&	0.914	&	0.914	&	0.914	&	0.914	&	0.597	&	0.889	&	0.8798	&	0.0946	&	0.914	&	0.914	\\
			K	&	71	&	1	&	1	&	1	&		1	&	1	&	1	&	1	&	0.977	&	0.977	&	0.977	&	0.9931	&	0.0105	&	1	&	1	\\
			L	&	66	&	0.668	&	0.668	&	0.668	&		0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0.668	&	0	&	0.668	&	0.668	\\
			M	&	79	&	0.706	&	0.706	&	0.706	&		0.706	&	0.706	&	0.706	&	0.721	&	0.721	&	0.667	&	0.745	&	0.709	&	0.0184	&	0.692	&	0.692	\\
			N	&	120	&	0.945	&	0.945	&	0.945	&		0.945	&	0.945	&	0.945	&	0.812	&	0.812	&	0.812	&	0.823	&	0.8929	&	0.064	&	0.945	&	0.914	\\
			O	&	97	&	0.912	&	0.912	&	0.912	&		0.912	&	0.912	&	0.912	&	0.912	&	0.912	&	0.929	&	0.929	&	0.9154	&	0.0068	&	0.948	&	0.948	\\
			P	&	511	&	0.824	&	0.824	&	0.824	&		0.821	&	0.821	&	0.818	&	0.818	&	0.833	&	0.83	&	0.827	&	0.824	&	0.0046	&	0.831	&	0.828	\\
			Q	&	401	&	0.74	&	0.74	&	0.74	&		0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0.74	&	0	&	0.813	&	0.75	\\
			R	&	425	&	0.895	&	0.895	&	0.895	&		0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0.895	&	0	&	0.885	&	0.92	\\
			S	&	336	&	0.836	&	0.836	&	0.836	&		0.836	&	0.836	&	0.836	&	0.836	&	0.836	&	0.836	&	0.831	&	0.8355	&	0.0015	&	0.831	&	0.831	\\
			T	&	473	&	0.85	&	0.85	&	0.85	&		0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0.85	&	0	&	0.863	&	0.863	\\
			U	&	412	&	0.929	&	0.929	&	0.929	&		0.929	&	0.81	&	0.81	&	0.81	&	0.806	&	0.806	&	0.925	&	0.8683	&	0.06	&	0.929	&	0.843	\\
			V	&	75	&	0.591	&	0.591	&	0.591	&		0.591	&	0.591	&	0.591	&	0.591	&	0.591	&	0.591	&	0.901	&	0.622	&	0.093	&	0.591	&	0.591	\\
			W	&	174	&	0.891	&	0.891	&	0.891	&		0.891	&	0.891	&	0.823	&	0.823	&	0.823	&	0.823	&	0.823	&	0.857	&	0.034	&	0.842	&	0.823	\\
			X	&	154	&	0.894	&	0.894	&	0.894	&		0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0.894	&	0	&	0.894	&	0.894	\\

			\bottomrule
		\end{tabular}
	\end{adjustbox}
	\caption{MCC of predicted structures with \OurTool{} through 10 runs. The Average and the standard deviation (Sd) of the MCC over the 10 runs are reported alongside with the MCC of the MFE and MEA structures obtained with \Software{RNAfold}.}.
\end{table}

\Heatset{min=0.4,  
	max=1,   
	max colour=Aquamarine3, % colour at maximum
	min colour=BadCol,      % colour at minimum
	Min colour=BadCol, % colour for values below min
	Max colour=Aquamarine3   % colour for values above max
}
\begin{table}[H]
{\centering \begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
\begin{tabular}{llllHllH}
	\toprule
	      &        & \multicolumn{3}{l}{\Software{Rsample}}    & \multicolumn{3}{l}{\Software{IPANEMAP}}   \\
	{RNA} & Length & Sensitivity & PPV    & \multicolumn1c{GM} & Sensitivity & PPV    & \multicolumn1c{GM} \\ \midrule
	A     & 34     & 0.625       & 1      & 0.791              & 0.625       & 1      & 0.791              \\
	B     & 530    & 0.6622      & 0.6203 & 0.641              & 0.8986      & 0.8062 & 0.851              \\
	C     & 301    & 0.03        & 0.0313 & 0.031              & 0.6         & 0.6    & 0.6                \\
	D     & 214    & 0.6349      & 0.6667 & 0.651              & 0.7302      & 0.8519 & 0.789              \\
	E     & 500    & 0.6382      & 0.7823 & 0.707              & 0.5263      & 0.5517 & 0.539              \\
	F     & 47     & 0.4         & 0.75   & 0.548              & 0.6         & 0.75   & 0.671              \\
	G     & 118    & 0.7179      & 0.8    & 0.758              & 0.7692      & 0.8333 & 0.801              \\
	H     & 76     & 0.4762      & 0.5556 & 0.514              & 0.7619      & 0.7619 & 0.762              \\
	I     & 155    & 0.6786      & 0.8261 & 0.749              & 0.9821      & 0.9821 & 0.982              \\
	J     & 79     & 0.7727      & 0.85   & 0.81               & 0.9545      & 0.875  & 0.914              \\
	K     & 71     & 1           & 1      & 1                  & 1           & 0.9545 & 0.977              \\
	L     & 66     & 0.5625      & 0.6923 & 0.624              & 0.625       & 0.7143 & 0.668              \\
	M     & 79     & 0.6538      & 0.8095 & 0.727              & 0.6538      & 0.68   & 0.667              \\
	N     & 120    & 0.2857      & 0.2632 & 0.274              & 0.9714      & 0.9189 & 0.945              \\
	O     & 97     & 0.9286      & 0.9286 & 0.929              & 0.9286      & 0.8966 & 0.912              \\
	P     & 511    & 0.9328      & 0.7762 & 0.851              & 0.8908      & 0.7681 & 0.827              \\
	Q     & 401    & 0.5826      & 0.5583 & 0.57               & 0.7304      & 0.75   & 0.74               \\
	R     & 425    & 0.8397      & 0.8271 & 0.833              & 0.9084      & 0.8815 & 0.895              \\
	S     & 336    & 0.5481      & 0.6129 & 0.58               & 0.8077      & 0.866  & 0.836              \\
	T     & 473    & 0.7986      & 0.7143 & 0.755              & 0.8819      & 0.8194 & 0.85               \\
	U     & 412    & 0.5379      & 0.6174 & 0.576              & 0.75        & 0.8462 & 0.797              \\
	V     & 75     & 0.619       & 0.6842 & 0.651              & 0.619       & 0.5652 & 0.591              \\
	W     & 174    & 0.8095      & 0.8644 & 0.836              & 0.8095      & 0.9808 & 0.891              \\
	X     & 154    & 0.875       & 0.913  & 0.894              & 0.875       & 0.913  & 0.894              \\ \bottomrule
\end{tabular}
\end{adjustbox}\\}

\caption{\OurTool~ prediction performance compared to \Software{Rsample} tool}
\end{table}

%  \Heatset{min=0,  
%	max=1,   
%	max colour=Aquamarine3, % colour at maximum
%	min colour=BadCol,      % colour at minimum
%	Min colour=BadCol, % colour for values below min
%	Max colour=Aquamarine3   % colour for values above max
%}

%  \HeatSecset{min=0,   
%	max=.1,   
%	max colour=BadCol, % colour at maximum
%	min colour=Aquamarine3,      % colour at minimum
%	Min colour=Aquamarine3, % colour for values below min
%	Max colour=BadCol   % colour for values above max
%}

\newpage 
\section{Detailed results of Lariat capping ribozyme from {\itshape Didymium iridis}}

\begin{table}[H]
 \Heatset{min=0.2,  
	max=.5,   
	max colour=BadCol, % colour at maximum
	min colour=Aquamarine3,      % colour at minimum
	Min colour=Aquamarine3, % colour for values below min
	Max colour=BadCol   % colour for values above max
}
\begin{center}
\begin{tabular}{lHH}\toprule
 & \multicolumn{2}{c}{Distance of reactivity}\\
Condition & \multicolumn1c{to prediction}&\multicolumn1c{to native}\\
\toprule
\OneMSevILUMg & 0.234 & 0.2872\\
\OneMSevILU &	0.2784 & 0.2961\\
\NMIAMg &	0.2606 & 0.2518\\
\NMIA &	0.2961 & 0.3316\\
\OneMSevILUThreeMg &	0.4291 & 0.3936\\
\NMIAMgCE &	0.4291 & 0.3404\\
\NAIMg 	& 0.3493 & 0.305\\
\BzCNMg & 	0.3759 & 0.2429\\
\OneMSevMgCE &	0.4379 & 0.3759\\
\CMCTMg &	0.4911 & 0.4113\\
\OneMSevILUThree &	0.2961 & 0.3404\\
\DMSMg &	0.3493 & 0.3227\\
\NMIACE 	&0.2695 &0.2518 \\
\OneMSevCE &	0.3582 & 	0.3316\\
\BzCN &	0.4911 & 	0.4734\\
\NAICE &	0.4025 &0.4025\\
\bottomrule
\end{tabular}
\end {center}
\caption{Distance of the predicted structure from the considered probing data.}
\end{table}

\begin{table}[H]
\newcommand{\MyScale}{.18}
\begin{adjustbox}{max width=\textwidth,keepaspectratio}
\begin{tabular}{@{}c|cccc@{}}

%\toprule
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/native_structure}&
 \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILU_NMIAMg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7ILUMg_1M7ILU3Mg}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/1M7MgCE_NMIAMgCE}& \includegraphics[scale=\MyScale]{graphs/Supp_structures/NaiMg_CMCTMg}\\
 \midrule
Native&\OneMSevILU&\OneMSevILUMg&\OneMSevMgCE&\NAIMg\\
structure &+\NMIAMg&+\OneMSevILUThreeMg&+\NMIAMgCE&+\CMCTMg\\
\end{tabular}
\end{adjustbox}
\caption{Predicted structures with conditions belonging to the same cluster}
\end{table}

\begin{table}[H]
\newcommand{\MA}[1]{\multicolumn{1}{c}{\Acc{#1}}}
\newcommand{\Acc}[1]{{\bfseries #1}}
\begin{adjustbox}{max width=\textwidth,max totalheight=\textheight,keepaspectratio}
\begin{tabular}{lHHHHHHHHHHHHHHHHH}
	\toprule
	                              & \multicolumn{1}{l}{Mono} & \MA{1} & \MA{2} & \MA{3} & \MA{4} & \MA{5} & \MA{6} & \MA{7} & \MA{8} & \MA{9} & \MA{10} & \MA{11} & \MA{12} & \MA{13} & \MA{14} & \MA{15} & \MA{16} \\ \midrule
	%Mono                         &                          & 0.85   & 0.84   & 0.81   & 0.8    & 0.74   & 0.73   & 0.73   & 0.71   & 0.7    & 0.7     & 0.64    & 0.63    & 0.61    & 0.6     & 0.6     & 0.6     \\
	\Acc{1} -- \OneMSevILUMg      & 0.85                     & 0.85   & 0.84   & 0.8    & 0.82   & 0.82   & 0.82   & 0.75   & 0.79   & 0.78   & 0.79    & 0.79    & 0.85    & 0.61    & 0.79    & 0.79    & 0.85    \\
	\Acc{2} -- \OneMSevILU        & 0.84                     &        & 0.84   & 0.82   & 0.82   & 0.7    & 0.83   & 0.72   & 0.82   & 0.85   & 0.83    & 0.84    & 0.84    & 0.84    & 0.6     & 0.84    & 0.6     \\
	\Acc{3} -- \NMIAMg            & 0.81                     &        &        & 0.81   & 0.8    & 0.74   & 0.82   & 0.82   & 0.81   & 0.81   & 0.79    & 0.8     & 0.8     & 0.61    & 0.81    & 0.81    & 0.6     \\
	\Acc{4} -- \NMIA              & 0.8                      &        &        &        & 0.8    & 0.74   & 0.73   & 0.81   & 0.78   & 0.7    & 0.79    & 0.63    & 0.72    & 0.61    & 0.7     & 0.7     & 0.59    \\
	\Acc{5} -- \OneMSevILUThreeMg & 0.74                     &        &        &        &        & 0.74   & 0.74   & 0.7    & 0.74   & 0.74   & 0.7     & 0.63    & 0.74    & 0.74    & 0.74    & 0.74    & 0.74    \\
	\Acc{6} -- \NMIAMgCE          & 0.73                     &        &        &        &        &        & 0.73   & 0.73   & 0.73   & 0.73   & 0.73    & 0.61    & 0.73    & 0.61    & 0.73    & 0.73    & 0.6     \\
	\Acc{7} -- \NAIMg             & 0.73                     &        &        &        &        &        &        & 0.73   & 0.72   & 0.7    & 0.73    & 0.64    & 0.68    & 0.61    & 0.73    & 0.73    & 0.59    \\
	\Acc{8} -- \BzCNMg            & 0.71                     &        &        &        &        &        &        &        & 0.71   & 0.65   & 0.73    & 0.63    & 0.72    & 0.6     & 0.61    & 0.6     & 0.59    \\
	\Acc{9} -- \OneMSevMgCE       & 0.7                      &        &        &        &        &        &        &        &        & 0.7    & 0.7     & 0.7     & 0.7     & 0.61    & 0.61    & 0.6     & 0.6     \\
	\Acc{10} -- \CMCTMg           & 0.7                      &        &        &        &        &        &        &        &        &        & 0.7     & 0.63    & 0.74    & 0.61    & 0.6     & 0.6     & 0.6     \\
	\Acc{11} -- \OneMSevILUThree  & 0.64                     &        &        &        &        &        &        &        &        &        &         & 0.64    & 0.62    & 0.62    & 0.65    & 0.65    & 0.59    \\
	\Acc{12} -- \DMSMg            & 0.63                     &        &        &        &        &        &        &        &        &        &         &         & 0.63    & 0.61    & 0.6     & 0.6     & 0.6     \\
	\Acc{13} -- \NMIACE           & 0.61                     &        &        &        &        &        &        &        &        &        &         &         &         & 0.61    & 0.62    & 0.6     & 0.6     \\
	\Acc{14} -- \OneMSevCE        & 0.6                      &        &        &        &        &        &        &        &        &        &         &         &         &         & 0.6     & 0.61    & 0.59    \\
	\Acc{15} -- \BzCN             & 0.6                      &        &        &        &        &        &        &        &        &        &         &         &         &         &         & 0.6     & 0.6     \\
	\Acc{16} -- \NAICE            & 0.6                      &        &        &        &        &        &        &        &        &        &         &         &         &         &         &         & 0.6     \\ \bottomrule
\end{tabular}
\end{adjustbox}
\caption{MCC of structures prediction with \OurTool{} from any pair of conditions}.
\end{table}

%\begin{table}
%\begin{center}
%\begin{tabular}{ll}
%Sample size & MCC\\
%\toprule
%
%100 & 0.55 \\
%500 & 0.55  \\
%1000 &  0.81\\
%2000 & 	0.55  \\
%3000 &  0.77 \\
%5000 & 	0.55  \\ 
%10000& 0.8 \\
%\end{tabular}
%\end {center}
%\caption{ MCC of predicted structures in absence of probig data when varyingthe sample size }
%\end{table}

\begin{figure}[H]
%\includegraphics[width=\linewidth]{graphs/histog}
\includegraphics[width=\linewidth]{graphs/histog1}
\caption{Distribution of MCC values over 100 runs from Mutate-and-Map data}
\end{figure}

%\bibliographystyle{bioinformatics}
\bibliographystyle{nar}
%\bibliographystyle{natbib}%compa
\bibliography{biblio}

\end{document}